---
title: "02 — Exploratory Analysis"
author: "Rutuja Kadam"
output:
  word_document: default
  html_document: default
---

```{r setup, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
source("../R/helpers.R")

dir.create("../figures", showWarnings = FALSE, recursive = TRUE)
dir.create("../results", showWarnings = FALSE, recursive = TRUE)

data <- read_csv("../data/processed/anonymized_survey_data.csv", show_col_types = FALSE)
```

## Sample profile

```{r profile}
age_distribution <- data %>% count(age_group, sort = TRUE) %>% mutate(percent = n / sum(n))
occupation_distribution <- data %>% count(occupation, sort = TRUE) %>% mutate(percent = n / sum(n))
region_distribution <- data %>% count(region_type, sort = TRUE) %>% mutate(percent = n / sum(n))
education_distribution <- data %>% count(education, sort = TRUE) %>% mutate(percent = n / sum(n))

write_csv(age_distribution, "../results/age_distribution.csv")
write_csv(occupation_distribution, "../results/occupation_distribution.csv")
write_csv(region_distribution, "../results/region_distribution.csv")
write_csv(education_distribution, "../results/education_distribution.csv")

age_distribution
occupation_distribution
region_distribution
education_distribution
```

## Current product distribution

```{r products}
# Keep the product visualization focused on actual menstrual-hygiene products.
# Missing/not-applicable responses are documented separately rather than plotted as products.
excluded_product_responses <- data %>%
  filter(is.na(primary_product) | primary_product == "Do not need after menopause") %>%
  count(primary_product, .drop = FALSE, name = "n")

product_distribution <- data %>%
  filter(!is.na(primary_product), primary_product != "Do not need after menopause") %>%
  count(primary_product, sort = TRUE) %>%
  mutate(percent = n / sum(n))

write_csv(product_distribution, "../results/product_distribution.csv")
write_csv(excluded_product_responses, "../results/product_excluded_responses.csv")

product_distribution

product_distribution_plot <- ggplot(
  product_distribution,
  aes(x = fct_reorder(primary_product, n), y = n)
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Current primary menstrual-hygiene product",
    x = NULL,
    y = "Respondents"
  ) +
  theme_minimal()

product_distribution_plot

ggsave(
  "../figures/product_usage_distribution.png",
  plot = product_distribution_plot,
  width = 8,
  height = 5,
  dpi = 300
)
```

## Importance of product attributes

```{r attributes}
importance_cols <- c(
  "comfort_importance", "cost_importance", "availability_importance",
  "health_risk_importance", "leakage_protection_importance",
  "durability_importance", "reusability_importance", "variants_importance",
  "brand_importance", "quality_importance", "sustainability_importance",
  "affordability_importance"
)

attribute_summary <- data %>%
  summarise(across(all_of(importance_cols), ~mean(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "attribute", values_to = "mean_rating") %>%
  mutate(
    attribute_label = attribute %>%
      sub("_importance$", "", .) %>%
      gsub("_", " ", .) %>%
      tools::toTitleCase()
  ) %>%
  arrange(desc(mean_rating))

write_csv(attribute_summary, "../results/attribute_importance_summary.csv")
attribute_summary

attribute_summary_plot <- ggplot(
  attribute_summary,
  aes(x = fct_reorder(attribute_label, mean_rating), y = mean_rating)
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Average importance rating by product attribute",
    x = NULL,
    y = "Mean rating (1–5)"
  ) +
  theme_minimal()

attribute_summary_plot

ggsave(
  "../figures/product_attribute_importance.png",
  plot = attribute_summary_plot,
  width = 8,
  height = 5,
  dpi = 300
)
```

## Behavioral outcomes

```{r behavior}
changed_product_summary <- data %>% count(changed_product_5y) %>% mutate(percent = n / sum(n))
future_switch_summary <- data %>% count(future_switch_intention) %>% mutate(percent = n / sum(n))
alternative_usage_summary <- data %>% count(uses_alternatives) %>% mutate(percent = n / sum(n))

write_csv(changed_product_summary, "../results/changed_product_distribution.csv")
write_csv(future_switch_summary, "../results/future_switch_intention_distribution.csv")
write_csv(alternative_usage_summary, "../results/alternative_product_usage_distribution.csv")

changed_product_summary
future_switch_summary
alternative_usage_summary
```

## Cultural and social context

```{r culture}
discussion_comfort_summary <- data %>% count(discussion_comfort) %>% mutate(percent = n / sum(n))
cultural_taboos_summary <- data %>% count(cultural_taboos) %>% mutate(percent = n / sum(n))
social_stigma_summary <- data %>% count(social_stigma) %>% mutate(percent = n / sum(n))

write_csv(discussion_comfort_summary, "../results/discussion_comfort_distribution.csv")
write_csv(cultural_taboos_summary, "../results/cultural_taboos_distribution.csv")
write_csv(social_stigma_summary, "../results/social_stigma_distribution.csv")

discussion_comfort_summary
cultural_taboos_summary
social_stigma_summary
```

## Interpretation guardrail

The sample is a convenience sample and is concentrated among younger and student respondents. Descriptive patterns should therefore be interpreted as characteristics of this sample, not population estimates.
