---
title: "03 — Awareness Analysis"
author: "Rutuja Kadam"
output:
  word_document: default
  html_document: default
---

```{r setup, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(broom)
source("../R/helpers.R")

dir.create("../figures", showWarnings = FALSE, recursive = TRUE)
dir.create("../results", showWarnings = FALSE, recursive = TRUE)

data <- read_csv("../data/processed/anonymized_survey_data.csv", show_col_types = FALSE) %>%
  mutate(
    high_awareness = if_else(awareness_score > 0.70, 1L, 0L),
    region_type = factor(region_type),
    formal_menstrual_education = factor(
      formal_menstrual_education,
      levels = c("No", "Somewhat", "Yes")
    )
  )
```

## Awareness distribution

```{r distribution}
awareness_summary <- tibble(
  metric = c("n", "mean", "sd", "median", "min", "max"),
  value = c(
    sum(!is.na(data$awareness_score)),
    mean(data$awareness_score, na.rm = TRUE),
    sd(data$awareness_score, na.rm = TRUE),
    median(data$awareness_score, na.rm = TRUE),
    min(data$awareness_score, na.rm = TRUE),
    max(data$awareness_score, na.rm = TRUE)
  )
)

awareness_category_summary <- data %>%
  count(awareness_category) %>%
  mutate(percent = n / sum(n))

write_csv(awareness_summary, "../results/awareness_summary.csv")
write_csv(awareness_category_summary, "../results/awareness_category_distribution.csv")

awareness_summary
awareness_category_summary

awareness_plot <- ggplot(data, aes(x = awareness_score)) +
  geom_histogram(binwidth = 0.1, boundary = 0, closed = "left") +
  labs(
    title = "Distribution of awareness score",
    x = "Awareness score (0–1)",
    y = "Respondents"
  ) +
  theme_minimal()

awareness_plot

ggsave(
  "../figures/awareness_score_distribution.png",
  plot = awareness_plot,
  width = 8,
  height = 5,
  dpi = 300
)
```

## Awareness by formal education/training

```{r education}
awareness_by_education <- data %>%
  group_by(formal_menstrual_education) %>%
  summarise(
    n = n(),
    mean_awareness = mean(awareness_score, na.rm = TRUE),
    sd_awareness = sd(awareness_score, na.rm = TRUE),
    median_awareness = median(awareness_score, na.rm = TRUE),
    .groups = "drop"
  )

write_csv(awareness_by_education, "../results/awareness_by_education.csv")
awareness_by_education
```

## Group comparison

```{r group-test}
# Awareness scores are bounded and may not satisfy normality within every group.
# Kruskal-Wallis is therefore used as a robust group-comparison test.
awareness_group_test <- kruskal.test(
  awareness_score ~ formal_menstrual_education,
  data = data
)

awareness_group_test_tidy <- broom::tidy(awareness_group_test)
write_csv(awareness_group_test_tidy, "../results/awareness_group_test.csv")
awareness_group_test_tidy
```

## Parsimonious logistic model for high awareness

The original coursework specification included age, occupation, education, income, region, and formal menstrual education simultaneously. Several of those categorical variables contain very small groups, which produced quasi/complete-separation and convergence warnings with only 146 respondents.

For the portfolio analysis, the model is deliberately simplified to two interpretable predictors with adequate group sizes: **formal menstrual education** and **region type**. This reduces overfitting and makes the estimated odds ratios more defensible for an exploratory sample.

```{r logistic}
awareness_model <- glm(
  high_awareness ~ formal_menstrual_education + region_type,
  data = data,
  family = binomial()
)

awareness_model_results <- broom::tidy(
  awareness_model,
  exponentiate = TRUE,
  conf.int = TRUE
)

write_csv(awareness_model_results, "../results/awareness_logistic_model.csv")
awareness_model_results
```

## Model diagnostics

```{r diagnostics}
model_diagnostics <- broom::glance(awareness_model)

convergence_summary <- tibble(
  metric = c("converged", "fisher_scoring_iterations", "n_observations"),
  value = c(
    as.character(awareness_model$converged),
    as.character(awareness_model$iter),
    as.character(nobs(awareness_model))
  )
)

write_csv(model_diagnostics, "../results/awareness_logistic_model_diagnostics.csv")
write_csv(convergence_summary, "../results/awareness_logistic_convergence.csv")

summary(awareness_model)
model_diagnostics
convergence_summary
```

## Interpretation

This model is **explanatory and exploratory**, not a validated population-level prediction model. The survey is a small convenience sample, so estimates should be interpreted as associations within this sample. The simplified model is preferred over a larger specification because sparse demographic categories in the original model produced separation and unstable estimates.
