---
title: "04 — Behavioral Modeling"
author: "Rutuja Kadam"
output:
  word_document: default
  html_document: default
---

```{r setup, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(MASS)
library(broom)
source("../R/helpers.R")

dir.create("../results", showWarnings = FALSE, recursive = TRUE)

data <- read_csv("../data/processed/anonymized_survey_data.csv", show_col_types = FALSE) %>%
  mutate(
    changed_product_5y = factor(changed_product_5y, levels = c("No", "Yes")),
    future_switch_intention = factor(
      future_switch_intention,
      levels = c("No", "Maybe", "Yes"),
      ordered = TRUE
    ),
    discussion_comfort = factor(discussion_comfort),
    cultural_taboos = factor(cultural_taboos),
    social_stigma = factor(social_stigma),
    formal_menstrual_education = factor(formal_menstrual_education)
  )
```

## Why these outcomes?

Current product type is extremely imbalanced in this survey, so a primary-product classifier would provide a misleading impression of predictive performance. Two better-supported behavioral outcomes are used instead:

1. whether a respondent changed primary product during the previous five years;
2. ordered future switching intention: `No < Maybe < Yes`.

## Model 1 — Past product change

```{r changed-model}
change_model <- glm(
  changed_product_5y ~ satisfaction + uses_alternatives +
    cultural_influence + discussion_comfort + cultural_taboos +
    social_stigma + formal_menstrual_education,
  data = data,
  family = binomial()
)

change_model_results <- broom::tidy(
  change_model,
  exponentiate = TRUE,
  conf.int = TRUE
)

change_model_diagnostics <- broom::glance(change_model)

write_csv(change_model_results, "../results/product_change_logistic_model.csv")
write_csv(
  change_model_diagnostics,
  "../results/product_change_logistic_model_diagnostics.csv"
)

change_model_results
change_model_diagnostics
```

## Model 2 — Future switching intention

The initial ordinal specification included several sparse categorical predictors. In particular, `discussion_comfort` produced implausibly large coefficients and odds ratios, indicating quasi-separation rather than a credible behavioral effect.

The revised model therefore uses a smaller set of continuous/ordinal predictors with adequate variation across the sample. The goal is **stable explanatory modeling**, not maximizing the number of variables in the equation.

```{r switch-model}
switch_model <- MASS::polr(
  future_switch_intention ~ satisfaction + cultural_influence +
    cost_importance + quality_importance + sustainability_importance,
  data = data,
  Hess = TRUE
)

summary(switch_model)
```

## Odds ratios for the ordinal model

```{r switch-or}
coef_table <- coef(summary(switch_model))
terms <- rownames(coef_table)
coef_only <- coef_table[!grepl("\\|", terms), , drop = FALSE]

ordinal_or <- data.frame(
  term = rownames(coef_only),
  estimate = coef_only[, "Value"],
  std_error = coef_only[, "Std. Error"],
  row.names = NULL
) %>%
  mutate(
    z_value = estimate / std_error,
    p_value = 2 * pnorm(abs(z_value), lower.tail = FALSE),
    odds_ratio = exp(estimate),
    conf_low = exp(estimate - 1.96 * std_error),
    conf_high = exp(estimate + 1.96 * std_error)
  )

write_csv(ordinal_or, "../results/switching_ordinal_model.csv")
ordinal_or
```

## Ordinal model fit summary

```{r switch-fit}
ordinal_model_fit <- tibble(
  metric = c("AIC", "log_likelihood", "residual_deviance", "n_observations"),
  value = c(
    AIC(switch_model),
    as.numeric(logLik(switch_model)),
    switch_model$deviance,
    nobs(switch_model)
  )
)

write_csv(ordinal_model_fit, "../results/switching_ordinal_model_diagnostics.csv")
ordinal_model_fit
```

## Model-stability check

```{r switch-stability}
ordinal_stability <- ordinal_or %>%
  summarise(
    max_absolute_coefficient = max(abs(estimate), na.rm = TRUE),
    max_odds_ratio = max(odds_ratio, na.rm = TRUE),
    min_odds_ratio = min(odds_ratio, na.rm = TRUE),
    all_finite = all(is.finite(estimate)) && all(is.finite(odds_ratio))
  )

write_csv(ordinal_stability, "../results/switching_ordinal_model_stability.csv")
ordinal_stability
```

## Proportional-odds caveat

`polr` assumes proportional odds. In a publication or production setting, this assumption should be tested explicitly (for example with a Brant-style test or by comparing less restrictive ordinal/multinomial specifications). The model is retained because `No`, `Maybe`, and `Yes` form a defensible order and because the revised specification avoids the sparse categorical predictors that destabilized the original model.

## Interpretation guardrail

These are observational associations. They do not establish that satisfaction, cultural influence, cost, quality, sustainability, or other predictors *cause* product-switching intentions. The model is intended to demonstrate careful explanatory modeling on a small survey dataset.
